\bf The EPEC Model

### 2024-07-19 Documentation/EPEC_Model.tex
Latest version

\subsection{Neoclassical Viscosity Matrices}\label{vmatrix}
The $2\times 2$ dimensionless species-$a$ neoclassical viscosity matrix, $[\mu^{\,a}](r)$,  is defined to have the following elements:
$\mu_{00}^{\,a} = K_{00}^{\,a}$, $\mu_{01}^{\,a}= (5/2)\,K_{00}^{\,a}- K_{01}^{\,a}$,
$\mu_{10}^{\,a}= \mu_{01}^{\,a}$,
$\mu_{11}^{\,a} = K_{11}^{\,a} - 5\,K_{01}^{\,a}+(25/4)\,K_{00}^{\,a}$.
Here,
\begin{align}
 K_{jk}^{\,e} &= g_t\,\frac{4}{3\sqrt{\pi}}\int_0^\infty
\frac{e^{-x}\,x^{\,4+j+k}\,\nu_D^{\,e}(x)\,dx}{[x^{\,2}+\nu_{\ast\,e}\,\nu_D^{\,e}(x)]\,[x^{\,2}+(5\pi/8)\,(\omega_{t\,e}\,\tau_{ee})^{\,-1}\,\nu_T^{\,e}(x)]},\\[0.5ex]
\nu_D^{\,e}&= \frac{3\sqrt{\pi}}{4}\left[\left(1-\frac{1}{2\,x}\right)\psi(x)+\psi'(x)\right]+\frac{3\sqrt{\pi}}{4}\,Z_{\rm eff},\\[0.5ex]
\nu_\epsilon^{\,e}&= \frac{3\sqrt{\pi}}{2}\left[\psi(x)-\psi'(x)\right],\\[0.5ex]
\nu_T^{\,a}(x) &= 3\,\nu_D^{\,a}(x)+\nu_{\epsilon}^{\,a}(x),
\end{align}
and
\begin{align}
\psi(x) &= \frac{2}{\sqrt{\pi}}\int_0^x{\rm e}^{-t^{\,2}}\,dt -\frac{2}{\sqrt{\pi}}\,x\,{\rm e}^{-x^{\,2}},\\[0.5ex]
\psi'(x)&= \frac{2}{\sqrt{\pi}}\,x\,{\rm e}^{-x^{\,2}}.
\end{align}
Furthermore, 
\begin{align}
 K_{jk}^{\,i} &=g_t\,\frac{4}{3\sqrt{\pi}}\int_0^\infty
\frac{e^{-x}\,x^{\,2+j+k}\,\nu_D^{\,i}(x)\,dx}{[x+\nu_{\ast\,i}\,\nu_D^{\,i}(x)]\,[x+(5\pi/8)\,(\omega_{t\,i}\,\tau_{ii})^{\,-1}\,\nu_T^{\,i}(x)]},\\[0.5ex]
 \nu_D^{\,i}&= \frac{3\sqrt{\pi}}{4}\left[\left(1-\frac{1}{2\,x}\right)\psi(x)+\psi'(x)\right]\frac{1}{x}\nonumber\\[0.5ex]\phantom{===}
&\phantom{=}+\frac{3\sqrt{\pi}}{4}\,\alpha_I\left[\left(1-\frac{x_{iI}}{2\,x}\right)\psi\!\left(\frac{x}{x_{iI}}\right)
+\psi'\!\left(\frac{x}{x_{iI}}\right)\right]\frac{1}{x},\\[0.5ex]
 \nu_\epsilon^{\,i}&=\frac{3\sqrt{\pi}}{2}\left[\psi(x)-\psi'(x)\right]\frac{1}{x}\nonumber\\[0.5ex]\phantom{===}
&\phantom{=}+\frac{3\sqrt{\pi}}{2}\,\alpha_I\left[\frac{m_i}{m_I}\,\psi\!\left(\frac{x}{x_{iI}}\right)
-\psi'\!\left(\frac{x}{x_{iI}}\right)\right]\frac{1}{x},
\end{align}
and, finally, 
\begin{align}
 K_{jk}^{\,I} &= g_t\,\frac{4}{3\sqrt{\pi}}\int_0^\infty
\frac{e^{-x}\,x^{\,2+j+k}\,\nu_D^{\,I}(x)\,dx}{[x+\nu_{\ast\,I}\,\nu_D^{\,I}(x)]\,[x+(5\pi/8)\,(\omega_{t\,I}\,\tau_{II})^{\,-1}\,\nu_T^{\,I}(x)]},\\[0.5ex]
 \nu_D^{\,I}&= \frac{3\sqrt{\pi}}{4}\left[\left(1-\frac{1}{2\,x}\right)\psi(x)+\psi'(x)\right]\frac{1}{x}\nonumber\\[0.5ex]\phantom{===}
&\phantom{=}+\frac{3\sqrt{\pi}}{4}\,\frac{1}{\alpha_I}\left[\left(1-\frac{x_{Ii}}{2\,x}\right)\psi\!\left(\frac{x}{x_{Ii}}\right)
+\psi'\!\left(\frac{x}{x_{Ii}}\right)\right]\frac{1}{x},\\[0.5ex]
 \nu_\epsilon^{\,I}&= \frac{3\sqrt{\pi}}{2}\left[\psi(x)-\psi'(x)\right]\frac{1}{x}\nonumber\\[0.5ex]\phantom{===}
&\phantom{=}+\frac{3\sqrt{\pi}}{2}\,\frac{1}{\alpha_I}\left[\frac{m_I}{m_i}\,\psi\!\left(\frac{x}{x_{Ii}}\right)
-\psi'\!\left(\frac{x}{x_{Ii}}\right)\right]\frac{1}{x}.
\end{align}
Note that our expressions for the neoclassical viscosity
matrices interpolate in the most accurate manner possible
between the three standard neoclassical collisionality
regimes (i.e., the banana, plateau, and Pfirsch-Sch\"{u}ter
regimes).

@@

\section{Plasma Response in Inner Region}
\subsection{Linear Layer Widths}\label{linear}
Let
\begin{align}
\tau_R(r) &= \mu_0\,r^{\,2}\,\sigma_{ee}\,Q_{00}^{ee},\\[0.5ex]
\sigma_{ee}(r) &=\frac{n_e\,e^{\,2}\,\tau_{ee}}{m_e},\\[0.5ex]
\tau_H(r) &= \frac{R_0}{B_0\,g}\,\frac{\sqrt{\mu_0\,\rho}}{n\,s},\\[0.5ex]
\rho(r)&= m_i\left[n_i(r) + n_b(r)\right] + m_I\,n_I(r),\\[0.5ex]
\tau_E(r) &= \frac{r^{\,2}}{D_\perp + (2/3)\,\chi_e},\\[0.5ex]
\tau_M(r) &= \frac{R_0^{\,2}\,q^{\,2}}{\chi_\phi},\\[0.5ex]
\rho_s (r)&= \frac{\sqrt{m_i\,T_e}}{e\,B_0\,g},\\[0.5ex]
\tau(r)& = -\frac{\omega_{\ast\,i}}{\omega_{\ast\,e}},\\[0.5ex]
S(r) &= \frac{\tau_R}{\tau_H},\\[0.5ex]
P_M(r) &= \frac{\tau_R}{\tau_M},\\[0.5ex]
P_E(r) &= \frac{\tau_R}{\tau_E},\\[0.5ex]
D(r)&= \frac{5}{3}\,S^{\,1/3}\,\frac{\rho_s}{r},\\[0.5ex]
Q_E(r)&= S^{\,1/3}\,n\,\omega_E\,\tau_H,\\[0.5ex]
Q_{e,i}(r)&= -S^{\,1/3}\,n\,\omega_{\ast\,e,i}\,\tau_H
\end{align}
Here, $\chi_e(r)$, $D_\perp(r)$, and $\chi_\phi(r)$ are the perpendicular electron energy, particle,
and toroidal momentum diffusivity profiles, respectively. Moreover, $n_b(r)$ is the number density of fast (as opposed to thermal) majority ions. 
The constant-$\psi$ linear layer width is determined from the solution of
\begin{equation}\label{e87}
\frac{d^{\,2}\,Y}{dp^{\,2}} -\left[\frac{-Q_E\,(Q_E-Q_i)+{\rm i}\,(Q_E-Q_i)\,(P_M+P_E)\,p^{\,2}+
P_M\,P_E\,p^{\,4}}
{{\rm i}\,(Q_E-Q_e) +\{P_E + {\rm i}\,(Q_E-Q_i)\,D^{\,2}\}\,p^{\,2}+(1+\tau)\,P_M\,D^{\,2}\,p^{\,4}}
\right] p^{\,2}\, Y = 0.
\end{equation}
If the small-$p$ behavior of solution of the previous equation that is well-behaved as $p\rightarrow \infty$ is
\begin{equation}
Y(p) = Y_0\left[1 - c\,p+{\cal O}(p^{\,2})\right].
\end{equation}
then the 
linear layer width is 
\begin{equation}
\delta_{\rm linear}(r) = \frac{\pi\,|c|\,r}{S^{\,1/3}}.
\end{equation}
Note that Eq.~(\ref{e87}) can only be solved when $1+\tau>0$.

>>>>

\subsection{Neoclassical Viscosity Matrices}\label{vmatrix}
The $2\times 2$ dimensionless species-$a$ neoclassical viscosity matrix, $[\mu^{\,a}](r)$,  is defined to have the following elements:
$\mu_{00}^{\,a} = K_{00}^{\,a}$, $\mu_{01}^{\,a}= (5/2)\,K_{00}^{\,a}- K_{01}^{\,a}$,
$\mu_{10}^{\,a}= \mu_{01}^{\,a}$,
$\mu_{11}^{\,a} = K_{11}^{\,a} - 5\,K_{01}^{\,a}+(25/4)\,K_{00}^{\,a}$.
Here,
\begin{align}
 K_{jk}^{\,e} &= g_t\,\frac{4}{3\sqrt{\pi}}\int_0^\infty
\frac{e^{-x^{\,2}}\,x^{\,5+2\,(j+k)}\,\nu_D^{\,e}(x)\,dx}{[x^{\,4}+\nu_{\ast\,e}\,\nu_D^{\,e}(x)]\,[x^{\,4}+(5\pi/8)\,(\omega_{t\,e}\,\tau_{ee})^{\,-1}\,\nu_T^{\,e}(x)]},\\[0.5ex]
\nu_D^{\,e}&= \frac{3\sqrt{\pi}}{4}\left[\left(1-\frac{1}{2\,x^2}\right)\psi(x)+\psi'(x)\right]+\frac{3\sqrt{\pi}}{4}\,Z_{\rm eff},\\[0.5ex]
\nu_\epsilon^{\,e}&= \frac{3\sqrt{\pi}}{2}\left[\psi(x)-\psi'(x)\right],\\[0.5ex]
\nu_T^{\,a}(x) &= 3\,\nu_D^{\,a}(x)+\nu_{\epsilon}^{\,a}(x),
\end{align}
and
\begin{align}
\psi(x) &= \frac{2}{\sqrt{\pi}}\int_0^{x^{\,2}}\sqrt{t}\,{\rm e}^{-t}\,dt,\\[0.5ex]
\psi'(x)&= \frac{2}{\sqrt{\pi}}\,x\,{\rm e}^{-x^{\,2}}.
\end{align}
Furthermore, 
\begin{align}
 K_{jk}^{\,i} &=g_t\,\frac{4}{3\sqrt{\pi}}\int_0^\infty
\frac{e^{-x^{\,2}}\,x^{\,4+2\,(j+k)}\,\nu_D^{\,i}(x)\,dx}{[x^{\,3}+\nu_{\ast\,i}\,\nu_D^{\,i}(x)]\,[x^{\,3}+(5\pi/8)\,(\omega_{t\,i}\,\tau_{ii})^{\,-1}\,\nu_T^{\,i}(x)]},\\[0.5ex]
 \nu_D^{\,i}&= \frac{3\sqrt{\pi}}{4}\left[\left(1-\frac{1}{2\,x^{\,2}}\right)\psi(x)+\psi'(x)\right]\frac{1}{x}\nonumber\\[0.5ex]\phantom{===}
&\phantom{=}+\frac{3\sqrt{\pi}}{4}\,\alpha_I\left[\left(1-\frac{x_{iI}^{\,2}}{2\,x^{\,2}}\right)\psi\!\left(\frac{x}{x_{iI}}\right)
+\psi'\!\left(\frac{x}{x_{iI}}\right)\right]\frac{1}{x},\\[0.5ex]
 \nu_\epsilon^{\,i}&=\frac{3\sqrt{\pi}}{2}\left[\psi(x)-\psi'(x)\right]\frac{1}{x}\nonumber\\[0.5ex]\phantom{===}
&\phantom{=}+\frac{3\sqrt{\pi}}{2}\,\alpha_I\left[\frac{m_i}{m_I}\,\psi\!\left(\frac{x}{x_{iI}}\right)
-\psi'\!\left(\frac{x}{x_{iI}}\right)\right]\frac{1}{x},
\end{align}
and, finally, 
\begin{align}
 K_{jk}^{\,I} &= g_t\,\frac{4}{3\sqrt{\pi}}\int_0^\infty
\frac{e^{-x^{\,2}}\,x^{\,4+2\,(j+k)}\,\nu_D^{\,I}(x)\,dx}{[x^{\,3}+\nu_{\ast\,I}\,\nu_D^{\,I}(x)]\,[x^{\,3}+(5\pi/8)\,(\omega_{t\,I}\,\tau_{II})^{\,-1}\,\nu_T^{\,I}(x)]},\\[0.5ex]
 \nu_D^{\,I}&= \frac{3\sqrt{\pi}}{4}\left[\left(1-\frac{1}{2\,x^{\,2}}\right)\psi(x)+\psi'(x)\right]\frac{1}{x}\nonumber\\[0.5ex]\phantom{===}
&\phantom{=}+\frac{3\sqrt{\pi}}{4}\,\frac{1}{\alpha_I}\left[\left(1-\frac{x_{Ii}^{\,2}}{2\,x^{\,2}}\right)\psi\!\left(\frac{x}{x_{Ii}}\right)
+\psi'\!\left(\frac{x}{x_{Ii}}\right)\right]\frac{1}{x},\\[0.5ex]
 \nu_\epsilon^{\,I}&= \frac{3\sqrt{\pi}}{2}\left[\psi(x)-\psi'(x)\right]\frac{1}{x}\nonumber\\[0.5ex]\phantom{===}
&\phantom{=}+\frac{3\sqrt{\pi}}{2}\,\frac{1}{\alpha_I}\left[\frac{m_I}{m_i}\,\psi\!\left(\frac{x}{x_{Ii}}\right)
-\psi'\!\left(\frac{x}{x_{Ii}}\right)\right]\frac{1}{x}.
\end{align}
Note that our expressions for the neoclassical viscosity
matrices interpolate in the most accurate manner possible
between the three standard neoclassical collisionality
regimes (i.e., the banana, plateau, and Pfirsch-Sch\"{u}ter
regimes).

@@

\subsection{Glasser-Greene-Johnson Data}
Let $\hat{B}=|{\bf B}|/B_0$, $\hat{\nabla}= R_0\,\nabla$, $d\psi_p/dr = f(r)/R_0$, and 
\begin{align}
J_1(r) &= \oint\frac{1}{\hat{B}}\,\frac{d{\mit\Theta}}{2\pi},\\[0.5ex]
J_2(r) &= \oint \hat{B}\,\frac{d{\mit\Theta}}{2\pi},\\[0.5ex]
J_3(r) &= \oint\frac{1}{\hat{B}^{\,3}}\,\frac{d{\mit\Theta}}{2\pi},\\[0.5ex]
J_4(r) &= \oint\frac{1}{\hat{B}\,|\hat{\nabla}\psi_p|^{\,2}}\,\frac{d{\mit\Theta}}{2\pi},\\[0.5ex]
J_5(r) &= \oint\frac{\hat{B}}{|\hat{\nabla}\psi_p|^{\,2}}\,\frac{d{\mit\Theta}}{2\pi},\\[0.5ex]
J_6(r) &= \oint\frac{1}{\hat{B}^{\,3}\,|\hat{\nabla}\psi_p|^{\,2}}\,\frac{d{\mit\Theta}}{2\pi}.
\end{align}
It follows that
\begin{align}
E(r) &= - \frac{dp/d\psi_p}{(dq/d\psi_p)^{\,2}}\,\frac{1}{\gamma}\left[\frac{d}{d\psi_p}\!\left(\frac{J_1}{\gamma}\right) - g\,\frac{dq}{d\psi_p}\,\frac{J_1}{J_2}\right]J_5,\\[0.5ex]
F(r) &= \frac{(dp/d\psi_p)^{\,2}}{(dq/d\psi_p)^{\,2}}\,\frac{1}{\gamma^{\,2}}\left[g^{\,2}\left(J_5\,J_6- J_4^{\,2}\right)+ J_5\,J_3\right],
\\[0.5ex]
H(r)&= \frac{dp/d\psi_p}{dq/d\psi_p}\,\frac{g}{\gamma}\left(J_4 - \frac{J_1\,J_5}{J_2}\right),
\end{align}
where $P(r)$ is the total plasma pressure. 
Finally,
\begin{equation}
D_R(r)= E + F + H^{\,2}.
\end{equation}

\section{Plasma Response in Inner Region}
\subsection{Linear Layer Widths}\label{linear}
Let
\begin{align}
\tau_R(r) &= \mu_0\,r^{\,2}\,\sigma_{ee}\,Q_{00}^{ee},\\[0.5ex]
\sigma_{ee}(r) &=\frac{n_e\,e^{\,2}\,\tau_{ee}}{m_e},\\[0.5ex]
\tau_H(r) &= \frac{R_0}{B_0\,g}\,\frac{\sqrt{\mu_0\,\rho}}{n\,s},\\[0.5ex]
\rho(r)&= m_i\left[n_i(r) + n_b(r)\right] + m_I\,n_I(r),\\[0.5ex]
\tau_\perp(r) &= \frac{r^{\,2}}{D_\perp},\\[0.5ex]
\tau_\varphi(r) &= \frac{r^{\,2}}{\chi_\phi},\\[0.5ex]
\tau(r)& = -\frac{\omega_{\ast\,e}}{\omega_{\ast\,i}},\\[0.5ex]
d_\beta (r)&= \frac{\sqrt{(5/3)\,m_i\,[T_e + (n_i/n_e)\,T_i + (n_I/n_e)\,T_I]}}{e\,B_0\,g},\\[0.5ex]
S(r) &= \frac{\tau_R}{\tau_H},\\[0.5ex]
P_\varphi(r) &= \frac{\tau_R}{\tau_\varphi},\\[0.5ex]
P_\perp(r) &= \frac{\tau_R}{\tau_\perp},\\[0.5ex]
D(r)&= S^{\,1/3}\,\left(\frac{\tau}{1+\tau}\right)^{1/2}\,\frac{d_\beta}{r},\\[0.5ex]
Q_E(r)&= -S^{\,1/3}\,n\,\omega_E\,\tau_H,\\[0.5ex]
Q_{e,i}(r)&= -S^{\,1/3}\,n\,\omega_{\ast\,e,i}\,\tau_H
\end{align}
Here, $\chi_e(r)$, $D_\perp(r)$, and $\chi_\phi(r)$ are the perpendicular electron energy, particle,
and toroidal momentum diffusivity profiles, respectively. Moreover, $n_b(r)$ is the number density of fast (as opposed to thermal) majority ions. 
The constant-$\psi$ linear layer width is determined from the solution of
\begin{equation}\label{e87}
\frac{d^{\,2}\,Y}{dp^{\,2}} -\left[\frac{-Q_E\,(Q_E+Q_i)+{\rm i}\,(Q_E+Q_i)\,(P_\varphi+P_\perp)\,p^{\,2}+
P_\varphi\,P_\perp\,p^{\,4}}
{{\rm i}\,(Q_E+Q_e) +\{P_\perp + {\rm i}\,(Q_E+Q_i)\,D^{\,2}\}\,p^{\,2}+(1+1/\tau)\,P_\varphi\,D^{\,2}\,p^{\,4}}
\right] p^{\,2}\, Y = 0.
\end{equation}
If the small-$p$ behavior of solution of the previous equation that is well-behaved as $p\rightarrow \infty$ is
\begin{equation}
Y(p) = Y_0\left[1 - c\,p+{\cal O}(p^{\,2})\right].
\end{equation}
then the 
linear layer width is 
\begin{equation}
\delta_{\rm linear}(r) = \frac{\pi\,|c|\,r}{S^{\,1/3}}.
\end{equation}
Note that Eq.~(\ref{e87}) can only be solved when $1+1/\tau>0$.
